---
title: "Task 3"
author: "BDSI Cancer Data Science"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(forcats)
# library(DIMPLE)
```


## Loading Data

```{r}
ovarianData <- readRDS(
  "/Users/derekyao/Documents/BDSI/Cancer Data Science/Data/ovarianData.RDS"
)
lungData <- readRDS(
  "/Users/derekyao/Documents/BDSI/Cancer Data Science/Data/lungData.RDS"
)
```

Visualize the proportion of tumor cells in each image stratified by patient
```{r}
ggplot(
  lungData$lung_cells %>% group_by(
    slide_id, image_id
  ) %>% 
    dplyr::summarize(Proportion = sum(pheno == "CK", na.rm = T)/n()) ,
  aes(x = fct_reorder(slide_id, Proportion, .desc = T), y = Proportion)
) + 
  geom_boxplot() +
  geom_boxplot(inherit.aes = F, aes(y = Proportion), col = "red") +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  labs(x = "Slide ID", title = "Proportion of Tumor Cells by Patient")

## the red line shows the distribution of the entire data set

ggplot(
  lungData$lung_cells %>% group_by(
    slide_id, pheno
  ) %>% 
    dplyr::summarize(total = n()) %>% 
    group_by(slide_id) %>% 
    mutate(Proportion = total / sum(total)),
  aes(y = Proportion, x = pheno)
) + geom_boxplot() +
  labs(x = "Phenotype")

ggplot(
  ovarianData$ovarian_cells %>% group_by(
    sample_id, pheno
  ) %>% 
    dplyr::summarize(total = n()) %>% 
    group_by(sample_id) %>% 
    mutate(Proportion = total / sum(total)),
  aes(y = Proportion, x = pheno)
) + geom_boxplot() +
  labs(x = "Phenotype")
  
```

Visualize the proportion of cells of each type, stratified by stage at imaging

```{r}
ggplot(
  lungData$lung_cells %>% group_by(
    slide_id, pheno
  ) %>% dplyr::summarize(
    total = n()
  ) %>% group_by(
    slide_id
  ) %>% 
    mutate(
      prop = total/sum(total)
    ) %>% left_join(
      lungData$lung_metadata[ , c("slide_id" , "stage_at_diagnosis")], by = "slide_id"
    ), aes(col = stage_at_diagnosis , y = prop, x = pheno)
) +
  geom_boxplot() +
  labs(x = "Phenotype", y = "Patient Prevalence", col = "Stage", title = "Visualizing Patient-Level Prevalence")
```

## Exploratory Analyses

```{r, results=FALSE}
propMetadata_lung <- left_join(
  lungData$lung_metadata,
  lungData$lung_cells %>% group_by(
    slide_id, pheno
  ) %>% dplyr::summarize(
    total = n()
  ) %>% group_by(
    slide_id
  ) %>% 
    mutate(
      prop = total/sum(total)
    ) %>% dplyr::ungroup() %>% dplyr::select(slide_id, pheno, prop) %>% 
  as.data.frame() %>% reshape(
      direction = "wide", timevar = "pheno", idvar = "slide_id"
    ), by = "slide_id"
)

rowSums(
  propMetadata_lung[
     , c("prop.CD14", "prop.CD19", "prop.CD4", "prop.CD8", "prop.CK", "prop.Other", "prop.NA")
  ]
)

propMetadata_lung[
   , c("prop.CD14", "prop.CD19", "prop.CD4", "prop.CD8", "prop.CK", "prop.Other", "prop.NA")
][is.na(
  propMetadata_lung[
     , c("prop.CD14", "prop.CD19", "prop.CD4", "prop.CD8", "prop.CK", "prop.Other", "prop.NA")
  ]
)] <- 0 # zeros occur because a phenotype is not present
# if not present, prevalence should be zero
rowSums(
  propMetadata_lung[
     , c("prop.CD14", "prop.CD19", "prop.CD4", "prop.CD8", "prop.CK", "prop.Other", "prop.NA")
  ]
)
# now prevalences add to 1

propMetadata_ovarian <- left_join(
  ovarianData$ovarian_metadata,
  ovarianData$ovarian_cells %>% group_by(
    sample_id, pheno
  ) %>% dplyr::summarize(
    total = n()
  ) %>% group_by(
    sample_id
  ) %>% 
    mutate(
      prop = total/sum(total)
    ) %>% dplyr::ungroup() %>% dplyr::select(sample_id, pheno, prop) %>% 
  as.data.frame() %>% reshape(
      direction = "wide", timevar = "pheno", idvar = "sample_id"
    ), by = "sample_id"
)

rowSums(
  propMetadata_ovarian[
     , c("prop.B Cell", "prop.Cytotoxic T", "prop.Macrophage", "prop.Other", "prop.Tumor", "prop.T Helper")
  ]
)

propMetadata_ovarian[
     , c("prop.B Cell", "prop.Cytotoxic T", "prop.Macrophage", "prop.Other", "prop.Tumor", "prop.T Helper")
  ][is.na(
  propMetadata_ovarian[
     , c("prop.B Cell", "prop.Cytotoxic T", "prop.Macrophage", "prop.Other", "prop.Tumor", "prop.T Helper")
  ]
)] <- 0 # zeros occur because a phenotype is not present
# if not present, prevalence should be zero
rowSums(
  propMetadata_ovarian[
     , c("prop.B Cell", "prop.Cytotoxic T", "prop.Macrophage", "prop.Other", "prop.Tumor", "prop.T Helper")
  ]
)
# now prevalences add to 1
```

```{r}
cor(
  propMetadata_lung$survival_days,
  propMetadata_lung[
     , c("prop.CD14", "prop.CD19", "prop.CD4", "prop.CD8", "prop.CK", "prop.Other", "prop.NA")
  ]
)
cor(
  propMetadata_ovarian$survival_time,
  propMetadata_ovarian[
     , c("prop.B Cell", "prop.Cytotoxic T", "prop.Macrophage", "prop.Other", "prop.Tumor", "prop.T Helper")
  ]
)

propMetadata_lung$survLong <- ifelse(
  propMetadata_lung$survival_days > median(propMetadata_lung$survival_days), 1, 0
)
propMetadata_ovarian$survLong <- ifelse(
  propMetadata_ovarian$survival_time > median(propMetadata_ovarian$survival_time), 1, 0
)

ggplot(
  propMetadata_lung %>% reshape2::melt(
    id.vars = c("slide_id", "survLong"), 
    measure.vars = c(
      "prop.CD14", "prop.CD19", "prop.CD4", "prop.CD8", "prop.CK", "prop.Other", "prop.NA"
    )
  ), aes(x = variable, y = value, col = as.factor(survLong))
) + geom_boxplot()
ggplot(
  propMetadata_ovarian %>% reshape2::melt(
    id.vars = c("sample_id", "survLong"), 
    measure.vars = c(
      "prop.B Cell", "prop.Cytotoxic T", "prop.Macrophage", "prop.Other", "prop.Tumor", "prop.T Helper"
    )
  ), aes(x = variable, y = value, col = as.factor(survLong))
) + geom_boxplot()
```
It looks like there could be a slight negative association between tumor cells
and median survival in the lung dataset. In the ovarian dataset, there is no 
association between survival and tumor cell prevalence. However, T Helper cell
prevalence may be associated with survival.